---
title: "data_wrangling"
format: html
---
Load the libraries 
```{r}
library(here)
library(janitor)
library(tidyverse)
library(ggplot2)
```

Download the data 
```{r}
stomach <- read_csv(here::here("data", "wg_stomachs.csv")) %>% 
  clean_names()

grunts <- read_csv(here::here("data", "wg.csv")) %>% 
  clean_names()

```

Data Wrangling 
```{r}
# Look at stomach datatypes 
str(stomach)

# NOTICE: stomach's `invert_parts` and `fish` are read as characters.
# Need to switch the data types to numeric! 
stomach$invert_parts <- as.numeric(stomach$invert_parts, na.rm = TRUE)
stomach$fish <- as.numeric(stomach$fish, na.rm = TRUE)



# Look at grunt datatypes 
str(grunts)

# NOTICE: grunt's `sl_cm`, `fl_cm`, `weight` are read as characters.
# Need to switch the data types to numeric! 
grunts$sl_cm <- as.numeric(grunts$sl_cm, na.rm = TRUE)
grunts$fl_cm <- as.numeric(grunts$fl_cm, na.rm = TRUE)
grunts$weight <- as.numeric(grunts$weight, na.rm = TRUE)

# Switch `date` be read as a date-time  
grunts$date <- as_date(grunts$date)

```


Add `total_gram` column (column with the total grams of each fish)
```{r}
stomach <- stomach %>% 
  rowwise() %>% # By Row
  mutate(total_grams = sum(c_across(5:14), na.rm = TRUE)) # Calculate the sum of all measurements across each row 

# `Mutate` makes a new row called 'total_grams'
```

Make each cell a percentage (gram measurement / total_grams)
```{r}
# Make a new dataframe called `stomach_perc` with `stomach` data
stomach_perc <- stomach 

# Divide all numeric cells by the total gram column 
stomach_perc[, 5:14] <- stomach_perc[, 5:14] / stomach_perc$total_grams  

```

Join stomach content and white grunt characteristic together 
```{r}
white_grunt_df <- left_join(stomach_perc, grunts, by = 'id_number') %>% # Join grunt data to stomach_perc data 
  select(-c(x19, snake, species.y, x11)) # removing unnecessary columns 

```


Exploring the data 
Looking a the mean percentages for each column 
```{r}
colMeans(white_grunt_df[, 5:12], na.rm = TRUE)
```

###Isolate the smallest and largest fish
library(dplyr)

# Smallest TL fish
smallest_fish <- grunts %>%
  filter(!is.na(tl_cm)) %>%
  slice_min(order_by = tl_cm, n = 1, with_ties = TRUE)

# Largest TL fish 
largest_fish <- grunts %>%
  filter(!is.na(tl_cm)) %>%
  slice_max(order_by = tl_cm, n = 1, with_ties = TRUE)

smallest_fish
largest_fish

###Error in the data entry...second smallest fish needs to be identified
library(dplyr)

second_smallest_fish <- grunts %>%
  filter(!is.na(tl_cm)) %>%
  distinct(tl_cm, .keep_all = TRUE) %>%  # keep unique TL values
  arrange(tl_cm) %>%
  slice(2)

second_smallest_fish

###Create pie chart with stomach content proportions

ggplot(diet_df, aes(x = "", y = proportion, fill = prey)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y") +
  labs(
    title = "Dietary Composition of White Grunt",
    fill = "Prey Category"
  ) +
  theme_void()

colMeans(white_grunt_df[, 5:12], na.rm = TRUE)


diet_means <- colMeans(white_grunt_df[, 5:12], na.rm = TRUE)


diet_df <- data.frame(
  prey = names(diet_means),
  proportion = diet_means
)


diet_df$percent <- diet_df$proportion * 100

library(ggplot2)

ggplot(diet_df, aes(x = "", y = percent, fill = prey)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y") +
  geom_text(
    aes(label = paste0(round(percent, 1), "%")),
    position = position_stack(vjust = 0.5),
    size = 4
  ) +
  labs(
    title = "Dietary Composition of White Grunt",
    fill = "Prey Category"
  ) +
  theme_void()

diet_means <- colMeans(
  white_grunt_df %>% select(-unknown_mush) %>% select(5:12),
  na.rm = TRUE
)



diet_df <- data.frame(
  prey = names(diet_means),
  percent = diet_means * 100
)

###Read otolith data
otolith_data <- read.csv("WG_Otolith_Data.csv")

